### 2020-06-02 writeup/proof_examples.tex
revision

\begin{proposition}[Labeled data efficiency]\label{prop_data_efficiency}
	In the isotropic model, assume that $\rho_1,\rho_2 > 10$ and $\Psi(\beta_1, \beta_2) < (5\rho_1)^{-1} + (5\rho_2)^{-1}$.
	Then the data efficiency ratio $x^\star$ satisfies 
	\[ x_l  \le x^\star\le \frac{1}{\rho_1 + \rho_2} \bigbrace{  \frac{2}{(\rho_1-1)^{-1} + (\rho_2-1)^{-1} - 5\Psi(\beta_1, \beta_2)}+1}, \]
	where we denoted
	$$x_l:= \frac1{\rho_1+\rho_2}\left(\frac{2}{(\rho_1-1)^{-1}+(\rho_2 -1)^{-1}}+1\right).$$
\end{proposition}

@@

Then we try to obtain an upper bound on $x^*$. In the following discussions, we only consider $x$ such that $x\ge \al_l$. In particular, we have $x(\rho_1+\rho_2) \ge x_l (\rho_1+\rho_2) \ge \min(\rho_1,\rho_2)$.

@@

\noindent{\bf The upper bound.} From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $x^*\le x$ if $x$ satisfies {\cor to revise}
\be\nonumber
\begin{split}
&(1+\oo(1)) \cdot \sum_{i=1}^2 pd^2 \frac{(\al\rho_i)^2\cdot \al (\rho_1+\rho_2)}{[\al ( \rho_1+\rho_2) - 1]^3}  \bigbrace{1 + \sqrt{\frac{1}{\al \rho_i}}}^4 \\
&\le \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{\al (\rho_1+\rho_2) - 1} . 
\end{split}
\ee
We rewrite the inequality as
\be
\begin{split}\label{solval1}
  (1+\oo(1)) \cdot \frac{pd^2}{[1 - (\al \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{\al \rho}}}^4\le \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{\al \rho  - 1},
\end{split}
\ee
where we abbreviated $\rho:=\rho_1+\rho_2$.

In order to solve \eqref{solval1}, we now consider the case $\min(\rho_1,\rho_2)\ge 200$. With some basic calculations, one can show that in this case
$$ \frac{1}{[1 - (\al \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho_1+\rho_2}} + \sqrt{\frac{1}{\al \rho}}}^4 <  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + 0.32.$$
Thus the following inequality implies \eqref{solval1}:
\be\label{solval1add}
\begin{split}
&\left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + 0.32\right) pd^2 < \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{\al (\rho_1+\rho_2) - 1} .
\end{split}
\ee
In particular, if 
\be\nonumber
\begin{split}
&\left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + 0.32\right) pd^2 < \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{(\rho_1+\rho_2) - 1} ,
\end{split}
\ee
that is, we have positive transfer when using all the data, then we can solve from \eqref{solval1add} the following upper bound on $\al^*$:
\begin{align*}
\al^* &<  \frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  -  \left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + 0.32\right)\frac{pd^2}{\sigma^2}}+ 1\right] \\
& < \frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1}+\frac1{\rho_2}  - \left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + \frac13\right)\frac{pd^2}{\sigma^2}}+ 1\right] .
\end{align*}

\noindent{\bf The lower bound.} From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $\al^*\ge \al$ if $\al$ satisfies
\be\label{solval2}
\begin{split}
&(1-\oo(1)) \cdot  \frac{pd^2}{[1 - (\al \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} - \sqrt{\frac{1}{\al \rho}}}^4 \ge \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{\al \rho - 1} . 
\end{split}
\ee
We then follow similar arguments as the above proof for the upper bound.

In order to solve \eqref{solval2}, we consider the case $\min(\rho_1,\rho_2)\ge 200$. With some basic calculations, one can show that the sum on the left-hand side of \eqref{solval2} satisfies 
$$ \frac{1}{[1 - (\al \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho_1+\rho_2}} - \sqrt{\frac{1}{\al \rho}}}^4 >  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} -0.26 .$$
Thus the following inequality implies \eqref{solval2}:
\be\label{solval2add}
\begin{split}
&\left( \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} -0.26\right) pd^2 > \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{\al (\rho_1+\rho_2) - 1} .
\end{split}
\ee
There are two cases: if 
\be\nonumber
\begin{split}
&\left( \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} -0.26\right) pd^2 \ge \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1},
\end{split}
\ee
then we have negative transfer for all choice of $0\le \al \le 1$; otherwise, we can solve from \eqref{solval2add} the following lower bound on $\al^*$:
\begin{align*}
\al^* &>  \frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  - \left( \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} -0.26\right) \frac{pd^2}{\sigma^2}}+ 1\right] \\
& >\frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1}+\frac1{\rho_2}  - \left( \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} -\frac13\right) \frac{pd^2}{\sigma^2}}+ 1\right] .
\end{align*}
This concludes the proof.
\end{proof}

@@

% Finally, we prove Proposition \ref{prop_covariate}, which shows that $\te(\hat{\beta}^{\MTL})$ is minimized approximately when $M$ is a scalar matrix where there is no model bias and enough source data.

\begin{proof}[Proof of Proposition \ref{prop_covariate}]
Let 
$$M_0:=\argmin_{M\in \cal S_{\mu}}g(M).$$ 
We now calculate $g(M_0)$. With the same arguments as in Lemma \ref{lem_hat_v} we can show that \eqref{hatv_add1} holds. Moreover, if the parameters are chosen such that $p^{c}\kappa^2  \le \sigma^2 \le p^{1-c} \kappa^2$ (recall \eqref{choiceofpara}), we can simplify
\be \nonumber
\begin{split}
g(M_0)&=(1+\OO(p^{-\e}))\cdot \sigma^2  \bigtr{\Sigma_2(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  ,
\end{split}
\ee
with high probability for any constant $\e>0$. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $\lambda(M)\in [\mu_{\min},\mu_{\max}]$. We omit the details here.

Now using Lemma \ref{lem_cov_shift}, we can obtain that with high probability,
\begin{align}\label{gvar_extra}
g(M_0)= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1(M_0)\cdot M_0^\top M_0 + a_2(M_0)}\right) \cdot \left(1 +\OO(p^{-\e})\right).
\end{align}
From equation \eqref{eq_a12extra}, it is easy to obtain the following estimates on $ a_1(M)$ and $a_2(M)$ for any $M\in \cal S_\mu$:
\be\label{est_a12extra}
\frac{\rho_1-1}{\rho_1+\rho_2} < a_1(M)<  \frac{\rho_1+\rho_2-1}{\rho_1+\rho_2},\quad a_2(M) < \frac{\rho_2}{\rho_1+\rho_2}.
\ee
Inserting \eqref{est_a12extra} into \eqref{gvar_extra} and using $\lambda(M_0^\top M_0)\ge \mu_{\min}^2$, we obtain that with high probability,
\begin{align}\label{approximateteM}
\left(1+\frac{\rho_2}{(\rho_1-1)\mu_{\min}^2}\right)^{-1}\wt g(M_0) \cdot \left(1 - \OO(p^{-\e})\right) \le g(M_0) \le \wt g(M_0) \cdot \left(1 +\OO(p^{-\e})\right),
\end{align}
where
$$\wt g(M_0):=\frac{\sigma^2}{(\rho_1+\rho_2)a_1(M_0)}\cdot \frac1p\tr\left( \frac{1}{M_0^\top M_0}\right) .$$
%With these two bounds, we can easily conclude \eqref{approxteM}. 
%
%We have that the test error satisfies
%\be\label{approxteM}  te(M)\left(1 -  \frac{n_2}{n_1-p} \frac{1}{\lambda_p^2 + \frac{n_2}{n_1-p}}\right)  \le  \frac{\sigma^2}{n_1+n_2}\tr\left( \frac{1}{a_1M^\top M + a_2}\right) \le te(M),\ee
%where $\lambda_p$ is the smallest singular value of $p$ and
%$$te(M):= \frac{\sigma^2}{a_1(n_1+n_2)}\tr\left( \frac{1}{M^\top M}\right) .$$
%Moreover, for all $M$ satisfying \eqref{GMcons}, the minimum of $te(M)$ is attained when $M= a\id$.
By AM-GM inequality, we observe that 
$$\tr\left( \frac{1}{M^\top M}\right) = \sum_{i=1}^p\frac{1}{\lambda_i}$$
is minimized when $\lambda_1 = \cdots\lambda_p=\mu$ under the restriction $\prod_{i=1}^p\lambda_i\le \mu^p$. Hence we get that 
\be\label{AMGM}\wt g(M_0) \le \frac{\sigma^2}{\mu^2 (\rho_1+\rho_2)a_1(M_0)}.\ee

On the other hand, when $M=\mu \id$, applying Lemma \ref{lem_cov_shift} we can obtain that with high probability,
\begin{align}\label{gvar_extra2}
\begin{split}
g(\mu \id)&= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{\mu^2 a_1 (\mu\id) + a_2(\mu\id)}\right) \cdot \left(1 +\OO(p^{-\e})\right)\\
&\le \frac{\sigma^2}{\mu^2(\rho_1+\rho_2)a_1 (\mu\id)}.
\end{split}
\end{align}
Combining \eqref{est_a12extra}, \eqref{approximateteM}, \eqref{AMGM} and \eqref{gvar_extra2}, we conclude the proof.
%, we conclude that the sum $\sum_{i=1}^p\lambda_i^{-1}$ is smallest when $\lambda_1=\cdots=\lambda_p = a$.
\end{proof}

@@

we observe the following two phases as we increase $n_1 / p$ in Figure \ref{fig_covariate}.

@@

We provide a theoretical justification below.

\begin{proof}[Theoretical justification of Example \ref{ex_complement}]
We denote the test error as $\te_\lambda(\hat \beta_{t}^{\MTL})$ in the setting where $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$. Then equations in \eqref{eq_a12extra} become
\be\label{compleeq} a_1 + a_2 = 1 - \frac{1}{\rho_1 + \rho_2},  \ \ a_1 + \frac{1}{2(\rho_1 + \rho_2)}\cdot \bigbrace{\frac{a_1}{a_1 + \lambda^2 a_2} + \frac{a_1}{a_1 + \frac{a_2}{\lambda^2}}} = \frac{\rho_1}{\rho_1 + \rho_2}. \ee
%It's not hard to verify that there is only one valid solution $(a_1,a_2)$ to \eqref{compleeq}. 
After solving these, with \eqref{gvar_extra} we get that whp,
\be\label{testcomple}
 \te_\lambda(\hat \beta_{t}^{\MTL})= \frac{\sigma^2}{2(\rho_1 + \rho_2)}(1+\OO(p^{-\e}))\cdot g(\lambda) ,\quad g(\lambda): = \frac{1}{\frac{a_1}{\lambda^2} + a_2} + \frac{1}{a_1\lambda^2 + a_2}.\ee

@@

%Second, from Figure \ref{fig_te_complement} we observe that the complementary cases with $\lambda>1$ is better than the case without covariate shift (i.e. $M=\id$ case) when $n_1<n_2$. On the other hand, if we have enough source task data such that $n_1>n_2$, then it is always better to have no covariate shift.
We now study the behavior of $g$ as $\lambda$ changes. 
%This phenomenon can be also explained using our theory. 
We abbreviate $\gamma:=(\rho_1 + \rho_2)^{-1}$. Then with \eqref{compleeq}, we can rewrite
$$g(\lambda)= \frac{1}{\frac{a_1}{\lambda^2} + (1-\gamma - a_1)} + \frac{1}{a_1\lambda^2 + (1-\gamma - a_1)}.$$
We can compute that
\begin{align*}
g(\lambda) - g(1)&= \frac{ \lambda^2-1}{1-\gamma} a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
&= \frac{(\lambda^2-1)^2}{1-\gamma}  a_1\cdot  \frac{2a_1 - (1-\gamma) }{[-a_1(\lambda^2-1)+(1-\gamma)\lambda^2 ][a_1(\lambda^2-1) + (1-\gamma)]} .
\end{align*}
From this expressions, we observe the following behaviors.
\begin{itemize}
\item[(i)] If $n_1>n_2$, we have $a_1>(1-\gamma)/2$ (because $a_1>a_2$ as observed from the equation \eqref{compleeq}). Hence $g(\lambda)>g(1)$, which gives $\te_\lambda(\hat \beta_{t}^{\MTL})>\te_1(\hat \beta_{t}^{\MTL})$.

\item[(ii)] If $n_1< n_2$, we have $a_1< (1-\gamma)/2$. Hence $g(\lambda)< g(1)$, which gives $\te_\lambda(\hat \beta_{t}^{\MTL})<\te_1(\hat \beta_{t}^{\MTL})$.

\item[(iii)] If $n_1=n_2$, we have $g(\lambda)=g(1)=2/(1-\gamma)$, which explains why the curves in Figure  \ref{fig_model_shift_phasetrans} (c) all cross at the point $n_1=n_2$.
\end{itemize}
These justify the observations in Figure \ref{fig_model_shift_phasetrans} (c), 
\end{proof}
%\todo{add a transition example}

@@

Then we show Proposition \ref{prop_var_transition}, which gives a transition threshold with respect to the difference between the noise levels of the two tasks.

\begin{proposition}[Labeled data de-noising]\label{prop_var_transition}
	In the isotropic model, assume that $\rho_1 > 40$ and $\ex{\norm{\beta_1 - \beta_2}^2}< \frac{1}{2} {\sigma_2^2}  \cdot \Phi(\rho_1, \rho_2)$.
	We derive the following transition as a parameter of $\sigma_1^2$:
	\begin{itemize}
		\item If $\sigma_1^2 \le -2 \rho_1 \cdot p d^2+\left(1+ \frac34\rho_1 \Phi(\rho_1, \rho_2)\right)\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\sigma_1^2 > - \frac12\rho_1\cdot p d^2   +\left(1+ \frac32\rho_1\Phi(\rho_1, \rho_2)\right) \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}
As a corollary, if $\sigma_1^2 \le \sigma_2^2$, then we always get positive transfer.
The proof of Proposition \ref{prop_var_transition} is similar to Proposition \ref{prop_dist_transition}.
%The details can be found in Appendix \ref{app_proof_data}.

>>>>

\begin{proposition}[Labeled data efficiency]\label{prop_data_efficiency}
	In the isotropic model, assume that $\rho_1,\rho_2 \ge 9$ and $\Psi(\beta_1, \beta_2) < (5(\rho_1-1))^{-1} + (5(\rho_2-1))^{-1}$.
	Then the data efficiency ratio $x^\star$ satisfies 
	\be\label{eq_uplowx} x_l  \le x^\star\le \frac{1}{\rho_1 + \rho_2} \bigbrace{  \frac{2}{(\rho_1-1)^{-1} + (\rho_2-1)^{-1} - 5\Psi(\beta_1, \beta_2)}+1}, \ee
	where we denoted
	$$x_l:= \frac1{\rho_1+\rho_2}\left(\frac{2}{(\rho_1-1)^{-1}+(\rho_2 -1)^{-1}}+1\right).$$
\end{proposition}

@@

Then we try to obtain an upper bound on $x^*$. In the following discussions, we only consider $x$ such that $x> x_l$. In particular, we have $x\rho > x_l \rho \ge \min(\rho_1,\rho_2)$, where we abbreviated $\rho:=\rho_1+\rho_2$.

@@

\noindent{\bf The upper bound.} From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $x^*\le x$ if $x$ satisfies 
\be\nonumber
\begin{split}
&(1+\oo(1)) \cdot \sum_{i=1}^2 pd^2 \frac{(x\rho_i)^2\cdot x\rho }{(x\rho - 1)^3}  \bigbrace{1 + \sqrt{\frac{1}{x \rho_i}}}^4 \le \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{x\rho  - 1} . 
\end{split}
\ee
We rewrite the inequality as
\be
\begin{split}\label{solval1}
  (1+\oo(1)) \cdot \frac{\Psi(\beta_1,\beta_2)}{[1 - (x\rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{x\rho}}}^4\le \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2}{x \rho  - 1}.
\end{split}
\ee
With $x\rho\ge \min(\rho_1,\rho_2)>9$, we can get the simple bound
$$\frac{1+\oo(1)}{[1 - (x\rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{x\rho}}}^4 < 5. $$
Inserting it into \eqref{solval1}, we can solve for the upper bound in \eqref{eq_uplowx}.
%where we abbreviated $\rho:=\rho_1+\rho_2$.

We can get better bounds if the values of $\rho_1$ and $\rho_2$ increase. For example, if we consider the case $\min(\rho_1,\rho_2)\ge 100$, then with some basic calculations, one can show that in this case
$$ \frac{1}{[1 - (x \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} + \sqrt{\frac{1}{x \rho}}}^4 <  \frac{\rho_1^2 + \rho_2^2}{\rho^2} + 0.52.$$
Thus the following inequality implies \eqref{solval1}:
\be\nonumber %\label{solval1add}
\begin{split}
&\left(  \frac{\rho_1^2 + \rho_2^2}{\rho^2} + 0.52\right) \Psi(\beta_1,\beta_2) < \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2 }{x\rho - 1} ,
\end{split}
\ee
%In particular, if 
%\be\nonumber
%\begin{split}
%&\left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + 0.32\right) pd^2 < \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{(\rho_1+\rho_2) - 1} ,
%\end{split}
%\ee
%that is, we have positive transfer when using all the data, then 
from which we can solve for the following upper bound on $\al^*$:
\begin{align*}
\al^* &<  \frac1{\rho} \frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  -  \left(  \frac{\rho_1^2 + \rho_2^2}{\rho^2} + 0.52\right)\Psi(\beta_1,\beta_2)}+ \frac1\rho .
%\\
%& < \frac1{\rho_1+\rho_2}\left[\frac{2 }{\frac{1}{\rho_1}+\frac1{\rho_2}  - \left(  \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} + \frac13\right)\frac{pd^2}{\sigma^2}}+ 1\right] .
\end{align*}
Similarly, we can get a better lower bound. 
%\noindent{\bf The lower bound.} 
From \eqref{reduceproof1} and \eqref{reduceproof2}, we see that $x^*\ge x$ if $x$ satisfies
\be\label{solval2}
\begin{split} 
&(1-\oo(1)) \cdot  \frac{\Psi(\beta_1,\beta_2)}{[1 - (x \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} - \sqrt{\frac{1}{x \rho}}}^4 \ge \frac{1}{\rho_1 -1}+\frac{1}{\rho_2 -1} - \frac{2 }{x \rho - 1} . 
\end{split}
\ee
%We then follow similar arguments as the above proof for the upper bound.
Then in the case $\min(\rho_1,\rho_2)\ge 100$, with some basic calculations, one can show that the sum on the left-hand side of \eqref{solval2} satisfies 
$$ \frac{1}{[1 - (x \rho)^{-1}]^3} \sum_{i=1}^2 \bigbrace{\sqrt{\frac{\rho_i}{\rho}} - \sqrt{\frac{1}{x\rho}}}^4 >  \frac{\rho_1^2 + \rho_2^2}{\rho^2} -0.33 .$$
Thus the following inequality implies \eqref{solval2}:
\be\label{solval2add}
\begin{split}
&\left( \frac{\rho_1^2 + \rho_2^2}{\rho^2} -0.33\right) pd^2 > \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1} - \frac{2\sigma^2}{x\rho - 1} ,
\end{split}
\ee
%There are two cases: if 
%\be\nonumber
%\begin{split}
%&\left( \frac{\rho_1^2 + \rho_2^2}{(\rho_1+\rho_2)^2} -0.26\right) pd^2 \ge \frac{\sigma^2}{\rho_1 -1}+\frac{\sigma^2}{\rho_2 -1},
%\end{split}
%\ee
%then we have negative transfer for all choice of $0\le \al \le 1$; otherwise, 
from which we can solve for the following lower bound on $x^*$:
\begin{align*}
x^* &>  \frac1\rho \frac{2 }{\frac{1}{\rho_1-1}+\frac1{\rho_2-1}  - \left( \frac{\rho_1^2 + \rho_2^2}{\rho^2} -0.33\right) \Psi(\beta_1,\beta_2)}+ \frac1\rho .
\end{align*}
This gives a lower bound above $x_l$.
\end{proof}

@@

We now prove Proposition \ref{prop_covariate}, which shows that $\te(\hat{\beta}^{\MTL})$ is minimized approximately when $M$ is a scalar matrix where there is enough source data.

\begin{proof}[Proof of Proposition \ref{prop_covariate}]
Let 
$$M_0:=\argmin_{M\in \cal S_{\mu}}g(M).$$ 
We now calculate $g(M_0)$. With the same arguments as in Lemma \ref{lem_hat_v} we can show that \eqref{hatv_add1} holds. Moreover, if the parameters are chosen such that $p^{-1+c_0} \sigma^2 \le \kappa^2  \le p^{-\e_0-c_0}\sigma^2$ as in \eqref{choiceofpara}, we can simplify
\be \nonumber
\begin{split}
g(M_0)&=(1+\OO(p^{-\e}))\cdot \sigma^2  \bigtr{\Sigma_2(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  ,
\end{split}
\ee
with high probability for some constant $\e>0$. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $ \mu_{\min}\le \lambda_p(M)\le \lambda_1(M)\le \mu_{\max}$. We omit the details here.

Now using Lemma \ref{lem_cov_shift}, we obtain that with high probability,
\begin{align}\label{gvar_extra}
g(M_0)= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1(M_0)\cdot M_0^\top M_0 + a_2(M_0)}\right) \cdot \left(1 +\OO(p^{-\e})\right).
\end{align}
From equation \eqref{eq_a12extra}, it is easy to obtain the following estimates on $ a_1(M)$ and $a_2(M)$ for any $M\in \cal S_\mu$:
\be\label{est_a12extra}
\frac{\rho_1-1}{\rho_1+\rho_2} < a_1(M)<  \frac{\rho_1+\rho_2-1}{\rho_1+\rho_2},\quad a_2(M) < \frac{\rho_2}{\rho_1+\rho_2}.
\ee
Inserting \eqref{est_a12extra} into \eqref{gvar_extra} and using $\lambda(M_0^\top M_0)\ge \mu_{\min}^2$, we obtain that with high probability,
\begin{align}\label{approximateteM}
\left(1+\frac{\rho_2}{(\rho_1-1)\mu_{\min}^2}\right)^{-1}h(M_0) \cdot \left(1 - \OO(p^{-\e})\right) \le g(M_0) \le h(M_0) \cdot \left(1 +\OO(p^{-\e})\right),
\end{align}
where
$$h(M_0):=\frac{\sigma^2}{(\rho_1+\rho_2)a_1(M_0)}\cdot \frac1p\tr\left( \frac{1}{M_0^\top M_0}\right) .$$
%With these two bounds, we can easily conclude \eqref{approxteM}. 
%
%We have that the test error satisfies
%\be\label{approxteM}  te(M)\left(1 -  \frac{n_2}{n_1-p} \frac{1}{\lambda_p^2 + \frac{n_2}{n_1-p}}\right)  \le  \frac{\sigma^2}{n_1+n_2}\tr\left( \frac{1}{a_1M^\top M + a_2}\right) \le te(M),\ee
%where $\lambda_p$ is the smallest singular value of $p$ and
%$$te(M):= \frac{\sigma^2}{a_1(n_1+n_2)}\tr\left( \frac{1}{M^\top M}\right) .$$
%Moreover, for all $M$ satisfying \eqref{GMcons}, the minimum of $te(M)$ is attained when $M= a\id$.
By AM-GM inequality, we observe that 
$$\tr\left( \frac{1}{M^\top M}\right) = \sum_{i=1}^p\frac{1}{\lambda_i^2}$$
is minimized when $\lambda_1 = \cdots\lambda_p=\mu$ under the restriction $\prod_{i=1}^p\lambda_i\le \mu^p$. Hence we get that 
\be\label{AMGM} h(M_0) \le \frac{\sigma^2}{\mu^2 (\rho_1+\rho_2)a_1(M_0)}.\ee

On the other hand, when $M=\mu \id$, applying Lemma \ref{lem_cov_shift} we obtain that with high probability,
\begin{align}\label{gvar_extra2}
\begin{split}
g(\mu \id)&= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{\mu^2 a_1 (\mu\id) + a_2(\mu\id)}\right) \cdot \left(1 +\OO(p^{-\e})\right)\\
&\le \frac{\sigma^2}{\mu^2(\rho_1+\rho_2)a_1 (\mu\id)}.
\end{split}
\end{align}
Combining \eqref{est_a12extra}, \eqref{approximateteM}, \eqref{AMGM} and \eqref{gvar_extra2}, we conclude the proof.
%, we conclude that the sum $\sum_{i=1}^p\lambda_i^{-1}$ is smallest when $\lambda_1=\cdots=\lambda_p = a$.
\end{proof}

@@

In the simulation of Figure \ref{fig_covariate}, we observe the following two phases as  $n_1 / p$ increases.

@@

We provide a theoretical justification now.

\begin{proof}[Theoretical justification of Example \ref{ex_complement}]
We denote the test error as $\te(\hat \beta_{2}^{\MTL},\lambda)$ in the setting where $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$. Then equations in \eqref{eq_a12extra} become
\be\label{compleeq} a_1 + a_2 = 1 - \frac{1}{\rho_1 + \rho_2},  \ \ a_1 + \frac{1}{2(\rho_1 + \rho_2)}\cdot \bigbrace{\frac{a_1}{a_1 + \lambda^2 a_2} + \frac{a_1}{a_1 + \lambda^{-2} a_2}} = \frac{\rho_1}{\rho_1 + \rho_2}. \ee
%It's not hard to verify that there is only one valid solution $(a_1,a_2)$ to \eqref{compleeq}. 
After solving these, with \eqref{gvar_extra} we get that whp,
\be\label{testcomple}
 \te(\hat \beta_{2}^{\MTL},\lambda)= \frac{\sigma^2}{2(\rho_1 + \rho_2)}(1+\OO(p^{-\e}))\cdot f(\lambda) ,\quad f(\lambda): = \frac{1}{\lambda^{-2}{a_1} + a_2} + \frac{1}{\lambda^2a_1 + a_2}.\ee

@@

%Second, from Figure \ref{fig_te_complement} we observe that the complementary cases with $\lambda>1$ is better than the case without covariate shift (i.e. $M=\id$ case) when $n_1<n_2$. On the other hand, if we have enough source task data such that $n_1>n_2$, then it is always better to have no covariate shift.
We now study the behavior of $f$ as $\lambda$ changes. 
%This phenomenon can be also explained using our theory. 
We abbreviate $\gamma:=(\rho_1 + \rho_2)^{-1}$. Then with \eqref{compleeq}, we can rewrite
$$f(\lambda)= \frac{1}{\lambda^{-2}{a_1} + (1-\gamma - a_1)} + \frac{1}{\lambda^2a_1 + (1-\gamma - a_1)}.$$
Then we can compute that
\begin{align*}
f(\lambda) - f(1)&= \frac{ \lambda^2-1}{1-\gamma} a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
&= \frac{(\lambda^2-1)^2}{1-\gamma}  a_1\cdot  \frac{2a_1 - (1-\gamma) }{[-a_1(\lambda^2-1)+(1-\gamma)\lambda^2 ][a_1(\lambda^2-1) + (1-\gamma)]} .
\end{align*}
From this expressions, we observe the following behaviors.
\begin{itemize}
\item[(i)] If $n_1>n_2$, we have $a_1>(1-\gamma)/2$ (because $a_1>a_2$ as observed from the equation \eqref{compleeq}). Hence $f(\lambda)>f(1)$, which gives $\te(\hat \beta_{2}^{\MTL},\lambda)>\te(\hat \beta_{2}^{\MTL},1)$.

\item[(ii)] If $n_1< n_2$, we have $a_1< (1-\gamma)/2$. Hence $f(\lambda)< f(1)$, which gives $\te(\hat \beta_{2}^{\MTL},\lambda)<\te(\hat \beta_{2}^{\MTL},1)$.

\item[(iii)] If $n_1=n_2$, we have $f(\lambda)=f(1)=2/(1-\gamma)$, which means $\te(\hat \beta_{2}^{\MTL},\lambda)$ and $\te(\hat \beta_{2}^{\MTL},1)$ are roughly the same. %, which explains why the curves in Figure  \ref{fig_model_shift_phasetrans} (c) all cross at the point $n_1=n_2$.
\end{itemize}
This also partially justifies Proposition \ref{prop_covariate}. %Theorem  the observations in Figure \ref{fig_model_shift_phasetrans} (c), 
\end{proof}
%\todo{add a transition example}

@@

In this section, we consider the case where the two tasks may have difference variances $\sigma_1^2\ne \sigma_2^2$ for the noises. In particular, we show Proposition \ref{prop_var_transition}, which gives a transition threshold with respect to the difference between the noise levels of the two tasks.

\begin{proposition}%[Labeled data de-noising]
\label{prop_var_transition}
	In the isotropic model, assume that $\rho_1 > 40$ and $\ex{\norm{\beta_1 - \beta_2}^2}< \frac{1}{2} {\sigma_2^2}  \cdot \Phi(\rho_1, \rho_2)$.
	Then we have the following transition with respect to $\sigma_1^2$:
	\begin{itemize}
		\item If $\sigma_1^2 < - \gamma_+^{1/2} \rho_1 \cdot p d^2+\left(1+ \gamma_+^{-1/2}\rho_1 \Phi(\rho_1, \rho_2)\right)\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_2^{\MTL}) < \te(\hat{\beta}_2^{\STL})$.
		\item If $\sigma_1^2 > -\gamma_-^{1/2}\rho_1\cdot p d^2   +\left(1+ \gamma_-^{-1/2}\rho_1\Phi(\rho_1, \rho_2)\right) \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_2^{\MTL}) > \te(\hat{\beta}_2^{\STL})$.
	\end{itemize}
\end{proposition}
As a corollary, if $\sigma_1^2 \le \sigma_2^2$, then we always get positive transfer.
%The proof of Proposition \ref{prop_var_transition} is similar to Proposition \ref{prop_dist_transition}.
%The details can be found in Appendix \ref{app_proof_data}.
